---
title: "p8105_hw6_yb2591"
author: "Yige Bian (yb2591)"
date: "2023-12-01"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
library(viridis)
library(modelr)
library(purrr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1
In the following data, we create `city_state` variable, `if_solved` varaible to indicate whether the homicide is solved, omit cities "Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL",  limit `victim_race` with white or black, convert `victim_age` to numeric.
```{r}
homicide_data = read_csv("./homicide-data.csv") |>
  mutate(city_state = paste(city, state, sep = ", "),
         victim_age = as.numeric(victim_age),
         resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)) |>
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")) & (victim_race %in% c("White", "Black")))
```
```{r}
baltimore_df = homicide_data |>
  filter(city_state == "Baltimore, MD") |>
  select(resolution, victim_age, victim_sex, victim_race)

p1_fit_bal = glm(resolution ~ victim_age + victim_sex + victim_race, data = baltimore_df, family = binomial())

p1_fit_bal |>
  broom::tidy()|> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == c("victim_sexMale")) |> 
  select(term, OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```
```{r}
p1_cities_model = homicide_data |>
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |>
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(term, city_state, OR, OR_CI_lower, OR_CI_upper)

p1_cities_model |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

# Problem 2
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```
```{r}
boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}

cal_quantities = function(df){
  fit = lm(tmax ~ tmin + prcp, data = df)
  
  r_sq = pull(broom::glance(fit), r.squared)
  log_b1b2 = fit |>
    broom::tidy() |>
    filter(!(term %in% "(Intercept)")) |>
    pull(estimate) |>
    prod() |>
    log()
  
  return(tibble(
    r_squared = r_sq,
    log_b1b2 = log_b1b2))
}

boot_straps = 
  tibble(strap_number = 1:5000) |>
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(df = weather_df))
  ) |>
  mutate(quantities = map(strap_sample, cal_quantities)) |>
    unnest(quantities)
```

```{r}
boot_straps |>
  ggplot(aes(x = r_squared)) + 
  geom_density() +
  labs(x = "R-squared", y = "Frequency") +
  ggtitle("Distribution of R-Squared")
```

```{r}
boot_straps |>
  drop_na(log_b1b2) |>
  ggplot(aes(x = log_b1b2)) + 
  geom_density() +
  labs(x = "R-squared", y = "Frequency") +
  ggtitle("Distribution of Estimated log(β1*β2)")
```

The bootstrap estimated r square gives the range `r round(range(pull(boot_straps, r_squared)), 3)`, mean `r `round(mean(pull(boot_straps, r_squared)), 3)`, and the plot shows that most estimated values are around 0.92, which shows that the model has a good fit of the data.

The plot of $log(\hat{\beta}_1*\hat{\beta}_2)$ shows an obvious left skewing, with the range `r round(range(pull(boot_straps, log_b1b2)), 3)`, and most of the values were around -5.5. We noticed that there are na in the value of $log(\hat{\beta}_1*\hat{\beta}_2)$ due to the negative result of the product of $\hat{\beta}_1$ and $\hat{\beta}_2$.

```{r}
ci_rsquared <- quantile(pull(boot_straps, r_squared), c(0.025, 0.975))
ci_logb1b2 <- quantile(pull(boot_straps, log_b1b2), c(0.025, 0.975), na.rm = TRUE)

ci_rsquared
ci_logb1b2
```
The 95% CI for $\hat{r}^2$ is (0.89, 0.94), and 95% CI for $log(\hat{\beta}_1*\hat{\beta}_2)$ is [-9.06, -4.54]

# Problem 3
Firstly, we load and clean the data. All variables are converted to appropriate type and there is no missing value through checking.
```{r}
birthweight = read_csv("./birthweight.csv") |>
  mutate(
    babysex = case_match(babysex,
      1 ~ "male",
      2 ~ "female"
    ),
    frace = case_match(frace,
      1 ~ "White",
      2 ~ "Black",
      3 ~ "Asian",
      4 ~ "Puerto Rican",
      8 ~ "Other",
      9 ~ "Unknown"
    ),
    malform = case_match(malform,
      0 ~ "absent",
      1 ~ "present"
    ),
    mrace = case_match(mrace,
      1 ~ "White",
      2 ~ "Black",
      3 ~ "Asian",
      4 ~ "Puerto Rican",
      8 ~ "Other"))
```
```{r}
skimr::skim(birthweight)
```

Firstly, we supposed that the birthweight is associated with babay's health status and characteristics (babysex, bhead, blength, gaweeks, malform), mom's health status and characteristics (delwt, momage, ppmbi, smoken, wtgain), financial status (fincome)
```{r}
model_one = lm(bwt ~ babysex + bhead + blength + gaweeks + delwt + malform + momage + ppbmi + smoken + wtgain, data = birthweight)

model_one |>summary()
```
After fitting the first model, we observed that malform and wtgain is not statistical significant for the model, so we delete these two variables and fit a new model.

```{r}
model_two = lm(bwt ~ babysex + bhead + blength + gaweeks + delwt + momage + ppbmi + smoken, data = birthweight)

model_two |>summary()
```
Now, all variables are significantly associated with birthweight and we use this as our final model.
```{r}
birthweight |>
  add_residuals(model_two) |>
  add_predictions(model_two) |>
  arrange(pred) |>
  ggplot(
    aes(x = pred ,y = resid)
  )+
  geom_point()+
  labs(title = "Residuals vs Predictions", x = "Predictions", y = "Residuals")
```

The plot shows that the fitted values show slightly unstable predictions when the birthweight is too small. Most residuals shows random pattern around 0.

Now we use cross validation to evluate our model and other two models.
```{r}
cv_df = crossv_mc(birthweight, 10) |>
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))|>
  mutate(
    our_model =  map(train, \(df) lm(bwt ~ babysex + bhead + blength + gaweeks + delwt + momage + ppbmi + smoken, data = birthweight)),
    sample_model_one = map(train, \(df) lm(bwt ~ blength + gaweeks, data = birthweight)),
    sample_model_two = map(train, \(df) lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = birthweight))) |>
  mutate(
    rmse_our_model = map2_dbl(our_model, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_sample_model_one = map2_dbl(sample_model_one, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_sample_model_two = map2_dbl(sample_model_two, test, \(mod, df) rmse(model = mod, data = df)))
```

```{r}
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
  stat_summary(fun = "mean", color = "black")
```
From the plot, it is obvious that in general, our model shows lower rmse than two sample models, and our model also shows the lowest mean of rmse.
